### 2025-07-10 factor-analysis.qmd
20250710 - players with highest levels on latent factors

```{r}
dataMerged <- full_join(
  player_stats_seasonal_avgPerGame,
  nfl_nextGenStats_seasonal %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
  by = c("player_id" = "player_gsis_id","season"),
)
```

@@

```{r}
lavaan::lavPredict(efa3factor_fit)
```

@@

efa3factorModified_syntax <- '
 # EFA Factor Loadings
 efa("efa1")*f1 + 
 efa("efa1")*f2 + 
 efa("efa1")*f3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
 passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
 avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
 avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
 passer_rating + completion_percentage_above_expectation

@@

```{r}
lavaan::lavPredict(efa3factorModified_fit)
```

@@

@@

```{r}
lavaan::lavPredict(cfa1factor_fit)
```

@@

```{r}
lavaan::lavPredict(cfa2factor_fit)
```

@@

```{r}
lavaan::lavPredict(cfa3factor_fit)
```

@@

 # Variances
 F1 ~~ 1*F1
 F2 ~~ 1*F2
 #F3 ~~ 1*F3
'
```

@@

```{r}
lavaan::lavPredict(cfa3factorModified_fit)
```

@@

>>>>

```{r}
dataMerged <- full_join(
  player_stats_seasonal_avgPerGame %>% select(-any_of("week")),
  nfl_nextGenStats_seasonal %>% select(-any_of(c("player_display_name","completions","attempts","receptions","targets"))),
  by = c("player_id" = "player_gsis_id","season"),
)
```

@@

```{r}
efa3factor_factorScores <- lavaan::lavPredict(efa3factor_fit)

head(efa3factor_factorScores)
```

@@

efa3factorModified_syntax <- '
 # EFA Factor Loadings
 efa("efa1")*F1 + 
 efa("efa1")*F2 + 
 efa("efa1")*F3 =~ completionsPerGame + attemptsPerGame + passing_yardsPerGame + passing_tdsPerGame + 
 passing_air_yardsPerGame + passing_yards_after_catchPerGame + passing_first_downsPerGame + 
 avg_completed_air_yards + avg_intended_air_yards + aggressiveness + max_completed_air_distance + 
 avg_air_distance + max_air_distance + avg_air_yards_to_sticks + passing_cpoe + pass_comp_pct + 
 passer_rating + completion_percentage_above_expectation

@@

```{r}
efa3factorModified_factorScores <- lavaan::lavPredict(efa3factorModified_fit)

head(efa3factorModified_factorScores)
```

@@

Here are the variables that had a standardized loading greater than 0.3 on each of the factors:

```{r}
factor1vars <- c(
  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")

factor2vars <- c(
  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")

factor3vars <- c(
  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
```

```{r}
efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F1"]),]
```

```{r}
dataForFA[which.max(efa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
```

```{r}
efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F2"]),]
```

```{r}
dataForFA[which.max(efa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
```

```{r}
efa3factorModified_factorScores[which.max(efa3factorModified_factorScores[,"F3"]),]
```

```{r}
dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
```

@@

```{r}
cfa1factor_factorScores <- lavaan::lavPredict(cfa1factor_fit)

head(cfa1factor_factorScores)
```

@@

```{r}
cfa2factor_factorScores <- lavaan::lavPredict(cfa2factor_fit)

head(cfa2factor_factorScores)
```

@@

```{r}
cfa3factor_factorScores <- lavaan::lavPredict(cfa3factor_fit)

head(cfa3factor_factorScores)
```

@@

 # Variances
 F1 ~~ 1*F1
 F2 ~~ 1*F2
'
```

@@

```{r}
cfa3factorModified_factorScores <- lavaan::lavPredict(cfa3factorModified_fit)

head(cfa3factorModified_factorScores)
```

@@

Here are the variables that loaded onto each of the factors:

```{r}
factor1vars <- c(
  "completionsPerGame","attemptsPerGame","passing_yardsPerGame","passing_tdsPerGame",
  "passing_air_yardsPerGame","passing_yards_after_catchPerGame","passing_first_downsPerGame")

factor2vars <- c(
  "avg_completed_air_yards","avg_intended_air_yards","aggressiveness","max_completed_air_distance",
  "avg_air_distance","max_air_distance","avg_air_yards_to_sticks")

factor3vars <- c(
  "passing_cpoe","pass_comp_pct","passer_rating","completion_percentage_above_expectation")
```

```{r}
cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F1"]),]
```

```{r}
dataForFA[which.max(cfa3factorModified_factorScores[,"F1"]), c("player_display_name", "season", factor1vars)]
```

```{r}
cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F2"]),]
```

```{r}
dataForFA[which.max(cfa3factorModified_factorScores[,"F2"]), c("player_display_name", "season", factor2vars)]
```

```{r}
cfa3factorModified_factorScores[which.max(cfa3factorModified_factorScores[,"F3"]),]
```

```{r}
dataForFA[which.max(cfa3factorModified_factorScores[,"F3"]), c("player_display_name", "season", factor3vars)]
```
